---
title: "Assignment 3"
author: "Huang Yingshan, Kahlbacher Fabian, Wattenberg Yannick, Aggeler Samuel"
date: "11/25/2021"
output: html_document
---

# Task 2
```{r}
# Loading the packages.
library(RSiena)
library(sna)
library(jaccard)
# Importing functions that are contained in .r files and will be useful for 
# estimating the model and formatting the results 
source("printSiena.R")
source("siena07ToConverge.R")

# loading data
# Importing the adjacency matrices of the networks
f1 <- as.matrix(read.csv("Glasgow/f1.csv", header = FALSE))
f2 <- as.matrix(read.csv("Glasgow/f2.csv", header = FALSE))
f3 <- as.matrix(read.csv("Glasgow/f3.csv", header = FALSE))
# Reading the demographic, alcohol comsuption and log distance characteristics of pupils
attributes <- read.csv("Glasgow/demographic.csv", header = TRUE)
logdistance <- as.matrix(read.csv("Glasgow/logdistance.csv", header = FALSE))
alcohol <- as.matrix(read.csv("Glasgow/alcohol.csv", header = TRUE))
```

## 1)
### 1.1)
```{r}
# Compute Jacard index
n <- nrow(f1)
j_12 <- sum(f1*f2)/(n^2-sum(f1+f2==0))
j_23 <- sum(f2*f3)/(n^2-sum(f2+f3==0))
j_12
j_23
```
Jaccard index between wave 1 and wave 2 is `r j_12`, between wave 2 and wave 3 is `r j_23`. The Jaccard index is in each case higher than 0.3. Thus, it indicates that the data, for each wave, is sufficiently informative for applying SAOMs. 

### 1.2)
```{r}
friendship <- sienaDependent(array(c(f1, f2, f3), dim = c(n, n, 3)))
gender <- coCovar(attributes$gender)
age <- coCovar(attributes$age)
alcoholConsumption <- varCovar(alcohol)
logDistance <- coDyadCovar(logdistance)
mydata <- sienaDataCreate(
  friendship,
  gender,
  age,
  alcoholConsumption,
  logDistance
)

# Data description
print01Report(mydata, modelname = "glasgow_init")
```

```{r}
# Model specification
# Include basic endogenous and exogenous variables
myeff <- getEffects(mydata)
myeff <- includeEffects(myeff, transTrip, cycle3)
myeff <- includeEffects(myeff, egoX, altX, sameX, interaction1="gender")

# H1: Students tend to be friends with popular pupils
myeff <- includeEffects(myeff, inPopSqrt) # outPopSqrt

# H2: Students tend to be friends with pupils with similar alcohol consumption to their own
myeff <- includeEffects(myeff, sameX, interaction1 ="alcoholConsumption")

# H3: Students tend to be friends with students that live in the same neighborhood (living nearby)
myeff <- includeEffects(myeff, X, interaction1 = "logDistance")

# Estimate the model
myAlgorithm <- sienaAlgorithmCreate(
  projname="friends_res",
  nsub = 3, n3 = 3000, seed = 1908
)
model.ev <- siena07(
  myAlgorithm,
  data=mydata,
  effects=myeff,
  returnDeps = TRUE,
  useCluster = TRUE,
  nbrNodes = 3
)
```

```{r}
t_conv.max <- function(model) {
  sqrt(
  t(apply(model$sf, 2, mean)) %*%
  solve(cov(model$sf)) %*%
  apply(model$sf, 2, mean)
 )
}
```


t-conv_k
```{r}
printSiena(model.ev)
```
all good everything <= 0.1

```{r}
t_conv.max(model.ev)
```
not good 0.3297 (should be <= 0.2)

fit it again
```{r}
model.ev <- siena07(
  myAlgorithm,
  data=mydata,
  effects=myeff,
  returnDeps = TRUE,
  useCluster = TRUE,
  nbrNodes = 4,
  prevAns = model.ev
)
```

t-conv_k
```{r}
printSiena(model.ev)
```
all good everything <= 0.1

```{r}
t_conv.max(model.ev1)
```
good now, since <= 0.2


### 1.3)
```{r}
# Goodness of Fit
# Indegree distribution
gof1.id <- sienaGOF(model.ev,
  verbose = FALSE, varName = "friendship",
  IndegreeDistribution
)

# Outdegree distribution
gof1.od <- sienaGOF(model.ev,
  verbose = FALSE, varName = "friendship",
  OutdegreeDistribution
)

# Triad census
gof1.tc <- sienaGOF(model.ev,
  verbose = FALSE, varName = "friendship",
  TriadCensus
)

# Geodesic distance
GeodesicDistribution <- function(i, data, sims, period, groupName,
                                 varName, levls = c(1:5, Inf), cumulative = TRUE) {
  x <- networkExtraction(i, data, sims, period, groupName, varName)
  require(sna)
  a <- sna::geodist(symmetrize(x))$gdist
  if (cumulative) {
    gdi <- sapply(levls, function(i) {
      sum(a <= i)
    })
  }
  else {
    gdi <- sapply(levls, function(i) {
      sum(a == i)
    })
  }
  names(gdi) <- as.character(levls)
  gdi
}

gof1.gd <- sienaGOF(model.ev,
  verbose = FALSE, varName = "friendship",
  GeodesicDistribution
)
```



```{r}
plot(gof1.id)
plot(gof1.id, center = TRUE, scale = TRUE)
```
Fits pretty well for all number of in-degrees. 


```{r}
plot(gof1.od)
```
The out-degree does not fits well. The only out-degrees which fit well are 0 and 4. 


```{r}
plot(gof1.tc, center = TRUE, scale = TRUE)
```
- to be described ...

```{r}
plot(gof1.gd, center = TRUE, scale = TRUE)
```
- to be described ...

In conclusion, does not fit to good. ...


### 1.4 
```{r}
printSiena(model.ev)
```


H1: Students tend to be friends with popular pupils

indegree - popularity (sqrt): 
  - significantly different from 0 on the 5% significance level. 
  - negative:  means that .... --> H1 is not supported .. usually different .. 
  
  
H2: Students tend to be friends with pupils with similar alcohol
consumption to their own

same alcoholConsumption: 
 - significantly different from 0 on the 5% significance level.
 - positive: means that .... --> H2 is supported 
 
 
H3: Students tend to be friends with students that live in the same
neighborhood (living nearby)

logDistance:
 - significantly different from 0 on the 5% significance level.
 - negative: means that: the greater the distance the smaller the probability to make a tie. --> H3 is supportet. 


## 2)
```{r}

# hyp 4: Popular students tend to increase or maintain their level of alcohol consumption
myeff <- includeEffects(myeff, indeg, interaction1="alcoholConsumption")

# hyp 5: Students tend to adjust their alcohol consumption to that of their friends
# maybe its that effect: avAlt
```




### Questions
- how does the model fits better or is it already okay? less variables?
- parameter interpretation in general? meaning of theta does it depends on the effect ... it is different than before when we tried to estimate how the probability of a tie changes, ...
- parameter estimates can be interpreted only while holding all other parameters constant? 