---
title: "Assignment 3"
author: "Huang Yingshan, Kahlbacher Fabian, Wattenberg Yannick, Aggeler Samuel"
date: "11/25/2021"
output:
  pdf_document: default
  html_document: default
---

# Task 2
```{r}
# Loading the packages.
library(RSiena)
library(sna)
#library(jaccard)
# Importing functions that are contained in .r files and will be useful for 
# estimating the model and formatting the results 
source("printSiena.R")
source("siena07ToConverge.R")

# loading data
# Importing the adjacency matrices of the networks
f1 <- as.matrix(read.csv("Glasgow/f1.csv", header = FALSE))
f2 <- as.matrix(read.csv("Glasgow/f2.csv", header = FALSE))
f3 <- as.matrix(read.csv("Glasgow/f3.csv", header = FALSE))
# Reading the demographic, alcohol comsuption and log distance characteristics of pupils
attributes <- read.csv("Glasgow/demographic.csv", header = TRUE)
logdistance <- as.matrix(read.csv("Glasgow/logdistance.csv", header = FALSE))
alcohol <- as.matrix(read.csv("Glasgow/alcohol.csv", header = TRUE))
```

## 1)
### 1.1)
```{r}
# Compute Jacard index
n <- nrow(f1)
j_12 <- sum(f1*f2)/(n^2-sum(f1+f2==0))
j_23 <- sum(f2*f3)/(n^2-sum(f2+f3==0))
j_12
j_23
```
Jaccard index between wave 1 and wave 2 is `r j_12`, between wave 2 and wave 3 is `r j_23`. The Jaccard index is in each case higher than 0.3. Thus, it indicates that the data, for each wave, is sufficiently informative for applying SAOMs. 

### 1.2)
```{r}
friendship <- sienaDependent(array(c(f1, f2, f3), dim = c(n, n, 3)))
gender <- coCovar(attributes$gender)
age <- coCovar(attributes$age)
alcoholConsumption <- varCovar(alcohol)
logDistance <- coDyadCovar(logdistance)
mydata <- sienaDataCreate(
  friendship,
  gender,
  age,
  alcoholConsumption,
  logDistance
)

# Data description
print01Report(mydata, modelname = "glasgow_init")
```

```{r}
# Model specification

# Include basic endogenous variables
myeff <- getEffects(mydata)

myeff <- includeEffects(
  myeff,
  density,
  recip,
  transTrip,
  transTies,
  cycle3,
  balance,
  outActSqrt
)

# Include basic exogenous variables
myeff <- includeEffects(myeff, egoX, altX, sameX, interaction1="gender")

# H1: Students tend to be friends with popular pupils
myeff <- includeEffects(myeff, inPopSqrt) # outPopSqrt

# H2: Students tend to be friends with pupils with similar alcohol consumption to their own
myeff <- includeEffects(myeff, sameX, interaction1 ="alcoholConsumption")

# H3: Students tend to be friends with students that live in the same neighborhood (living nearby)
myeff <- includeEffects(myeff, X, interaction1 = "logDistance")
```

```{r}
# Estimate the model
myAlgorithm <- sienaAlgorithmCreate(
  projname="friends_res",
  nsub = 4, n3 = 3000, seed = 1908
)
model.ev <- siena07(
  myAlgorithm,
  data=mydata,
  effects=myeff,
  returnDeps = TRUE,
  useCluster = TRUE,
  nbrNodes = 8
)
```

```{r}
t_conv.max <- function(model) {
  sqrt(
  t(apply(model$sf, 2, mean)) %*%
  solve(cov(model$sf)) %*%
  apply(model$sf, 2, mean)
 )
}
```


t-conv_k
```{r}
printSiena(model.ev)
```
Not all t.conv are below 0.1 which indicates that the model is not converged

```{r}
t_conv.max(model.ev)
```
The t.conv.max value (2.052) is also indicates that the model is not converged as it is not below 0.2.

fit it again
```{r}
model.ev <- siena07(
  myAlgorithm,
  data=mydata,
  effects=myeff,
  returnDeps = TRUE,
  useCluster = TRUE,
  nbrNodes = 8,
  prevAns = model.ev
)
```

t-conv_k
```{r}
printSiena(model.ev)
```
All t.conv are below 0.1 which indicates that the model is converged for this measure .

```{r}
t_conv.max(model.ev)
```
The t.conv.max value (0.5655) is still not below 0.2 because the model is not yet converged.


fit it again
```{r}
model.ev <- siena07(
  myAlgorithm,
  data=mydata,
  effects=myeff,
  returnDeps = TRUE,
  useCluster = TRUE,
  nbrNodes = 8,
  prevAns = model.ev
)
```

```{r}
printSiena(model.ev)
```
All t.conv are below 0.1 which indicates that the model is converged for this measure .

```{r}
t_conv.max(model.ev)
```
The t.conv.max value (0.333) is still not below 0.2 because the model is not yet converged.
```{r}
model.ev <- siena07(
  myAlgorithm,
  data=mydata,
  effects=myeff,
  returnDeps = TRUE,
  useCluster = TRUE,
  nbrNodes = 8,
  prevAns = model.ev
)
```

```{r}
printSiena(model.ev)
```

All t.conv are below 0.1 which indicates that the model is converged.

```{r}
t_conv.max(model.ev)
```
The t.conv.max value (0.1195636) is below 0.2 indicating the model is converged.

### 1.3)
```{r}
# Goodness of Fit
# Indegree distribution
gof1.id <- sienaGOF(model.ev,
  verbose = FALSE, varName = "friendship",
  IndegreeDistribution
)

# Outdegree distribution
gof1.od <- sienaGOF(model.ev,
  verbose = FALSE, varName = "friendship",
  OutdegreeDistribution
)

# Triad census
gof1.tc <- sienaGOF(model.ev,
  verbose = FALSE, varName = "friendship",
  TriadCensus
)

# Geodesic distance
GeodesicDistribution <- function(i, data, sims, period, groupName,
                                 varName, levls = c(1:5, Inf), cumulative = TRUE) {
  x <- networkExtraction(i, data, sims, period, groupName, varName)
  require(sna)
  a <- sna::geodist(symmetrize(x))$gdist
  if (cumulative) {
    gdi <- sapply(levls, function(i) {
      sum(a <= i)
    })
  }
  else {
    gdi <- sapply(levls, function(i) {
      sum(a == i)
    })
  }
  names(gdi) <- as.character(levls)
  gdi
}

gof1.gd <- sienaGOF(model.ev,
  verbose = FALSE, varName = "friendship",
  GeodesicDistribution
)
```



```{r}
plot(gof1.id)
plot(gof1.id, center = TRUE, scale = TRUE)
```
All in-degrees from 0 to 8 are in the 95% confidence interval. Furthermore, the p-value is greater than 0.05 which indicates a good fit (as mentioned in the lecture).

```{r}
plot(gof1.od)
```
All out-degrees from 0 to 8 are in the 95% confidence interval except for 7 which is outside. Furthermore, the p-value is greater than 0.05 which indicates a good fit.


```{r}
plot(gof1.tc, center = TRUE, scale = TRUE)
```
GOF for triad census does not fit very well. The observed triad census for the following states 111U, 030C, 120C, 210 and 300 lies outside the 95%-confidence interval. Moreover, the p-value is equal to zero which indicates a poor fit. 



```{r}
plot(gof1.gd, center = TRUE, scale = TRUE)
```
The GOF for the Geodesic Distance does not fit well. Only the Geodesic Distance of one lies in the 95%-confidence interval the distances 2 to 5 are outside.
Moreover, the p-value of zero indicates a poor fit. 

To conclude, the in and out-degree fit well but the model does not accurately represents Triadcensus and Geodesic Distance. Thus, the significance of the assessments based on this model should be interpreted carefully. 

To fit out-degree well we tried to include different endogenous variables. The exogeneous variables were alway the same as specified in the code. Our different trials were: 
1. endogenous: transTrip, cycle3
2. endogenous: density, recip, transTrip, cycle3
3. engogenous: density, recip, transTrip, transTies, cycle3, balance, outActSqrt
The 3. combination of endogeneous variables worked well to fit out-degree. 

### 1.4 
```{r}
printSiena(model.ev)
```
The estimates of the coefficients and the corresponding standard errors are obtained from a model controlling for basic endogenous and exogenous
variables (all variables can be seen in the summary output above). 

H1: Students tend to be friends with popular pupils

The coefficient for "indegree - popularity (sqrt)" is significantly different from 0 on the 5% significance level. The coefficient is positive which means that actors are more likely to name popular (high in-degree) students as their friends. Hence, H1 is supported.
  
H2: Students tend to be friends with pupils with similar alcohol
consumption to their own

The coefficient for "same alcoholConsumption" is significantly different from 0 on the 5% significance level. The coefficient is positive which means that an actor is more likely to name students as their friends if they have the same level of alcohol consumption. Same level of alcohol consumption is a proxy for similar level of alcohol consumption hence, H2 is supported.
 
H3: Students tend to be friends with students that live in the same
neighborhood (living nearby)

The coefficient for "logDistance" is significantly different from 0 on the 5% significance level. The coefficient is negative which means that actors are less likely to make ties to students which are further away. Hence, H3 is supported.

## 2)

```{r}
friendship <- sienaDependent(array(c(f1, f2, f3), dim = c(n, n, 3)))
drinkbeh <- sienaDependent(alcohol, type = "behavior")

gender <- coCovar(attributes$gender)
age <- coCovar(attributes$age)
logDistance <- coDyadCovar(logdistance)

mydata2 <- sienaDataCreate(
  friendship,
  drinkbeh,
  gender,
  age,
  logDistance
)

mydata2
```


```{r}
# Data description
print01Report(mydata2, modelname = "glasgow_coEvol")
```
The Jaccard Index for the first wave (0.304) and second wave (0.351) are at least 0.3 thus, the data is informative.

```{r}
moran1 <- nacf(f1, alcohol[, 1], lag.max = 1, type = "moran",
neighborhood.type = "out", mode = "digraph")
moran2 <- nacf(f2, alcohol[, 2], lag.max = 1, type = "moran",
neighborhood.type = "out", mode = "digraph")
moran3 <- nacf(f3, alcohol[, 3], lag.max = 1, type = "moran",
neighborhood.type = "out", mode = "digraph")
autocorr <- rbind(moran1, moran2, moran3)
autocorr[, 2]
```
The Moran Index is about 0.3 in all waves which is low but indicates some behavioral structure in the network. 

## 2.1)
```{r}
# Model specification
# Include basic endogenous 
myeff <- getEffects(mydata2)

myeff <- includeEffects(
  myeff,
  density,
  recip,
  transTrip,
  transTies,
  cycle3,
  balance,
  outActSqrt
)

# Include basic exogenous variables
myeff <- includeEffects(myeff, egoX, altX, sameX, interaction1="gender")

# H1: Students tend to be friends with popular pupils
myeff <- includeEffects(myeff, inPopSqrt) # outPopSqrt

# H2: Students tend to be friends with pupils with similar alcohol consumption to their own
myeff <- includeEffects(myeff, sameX, interaction1 ="drinkbeh")

# H3: Students tend to be friends with students that live in the same neighborhood (living nearby)
myeff <- includeEffects(myeff, X, interaction1 = "logDistance")


# H4: Popular students tend to increase or maintain their level of alcohol consumption
myeff <- includeEffects(myeff, outdeg, indeg,
                        name = "drinkbeh", interaction1 = "friendship")

# H5:  Students tend to adjust their alcohol consumption to that of their friends
myeff <- includeEffects(myeff, avSim,
                        name = "drinkbeh", interaction1="friendship")
```
## 2.2)
```{r}
# Estimate the model
myAlgorithm <- sienaAlgorithmCreate(
  projname="friends_res",
  nsub = 4, n3 = 1000, seed = 1908
)

model.ev <- siena07(
  myAlgorithm,
  data=mydata2,
  effects=myeff,
  returnDeps = TRUE,
  useCluster = TRUE,
  nbrNodes = 8
)
```

```{r}
model.ev
```
All convergence t-ratios are below 0.1 but the overall is larger than 0.2 (2.0040).

```{r}
myAlgorithm <- sienaAlgorithmCreate(
  projname="friends_res",
  nsub = 4, n3 = 3000, seed = 1908
)

model.ev <- siena07(
  myAlgorithm,
  data=mydata2,
  effects=myeff,
  returnDeps = TRUE,
  useCluster = TRUE,
  nbrNodes = 8,
  prevAns = model.ev
)
```

```{r}
model.ev
```
All convergence t-ratios are below 0.1 but the overall is larger than 0.2 (0.3588).

```{r}
model.ev <- siena07(
  myAlgorithm,
  data=mydata2,
  effects=myeff,
  returnDeps = TRUE,
  useCluster = TRUE,
  nbrNodes = 8,
  prevAns = model.ev
)
```

```{r}
model.ev
```
All convergence t-ratios are below 0.1 and the overall is below 0.2 (0.1359). The model converged well.


```{r}
# Goodness of Fit
# Indegree distribution
gof1.id <- sienaGOF(model.ev,
  verbose = FALSE, varName = "friendship",
  IndegreeDistribution
)

# Outdegree distribution
gof1.od <- sienaGOF(model.ev,
  verbose = FALSE, varName = "friendship",
  OutdegreeDistribution
)

# Triad census
gof1.tc <- sienaGOF(model.ev,
  verbose = FALSE, varName = "friendship",
  TriadCensus
)

# Geodesic distance
gof1.gd <- sienaGOF(model.ev,
  verbose = FALSE, varName = "friendship",
  GeodesicDistribution
)
```


```{r}
plot(gof1.id)
```
All in-degrees from 0 to 8 are in the 95% confidence interval. Furthermore, the p-value is greater than 0.05 which indicates a good fit (as mentioned in the lecture).

```{r}
plot(gof1.od )
```
All out-degrees from 0 to 8 are in the 95% confidence interval except for 6 which is outside. Furthermore, the p-value is greater than 0.05 which indicates a good fit.

```{r}
plot(gof1.tc, center=TRUE, scale=TRUE)
```
GOF for triad census does not fit very well. The observed triad census for the following states 030C, 120C, 210 and 300 lies outside the 95%-confidence interval. Moreover, the p-value is equal to zero which indicates a poor fit.

```{r}
plot(gof1.gd, center=T, scale=T)
```
The GOF for the Geodesic Distance does not fit well. Only the Geodesic Distance of one lies in the 95%-confidence interval the distances 2 to 5 are outside. Moreover, the p-value of zero indicates a poor fit. 

To conclude, the in and out-degree fit well but the model does not accurately represents Triadcensus and Geodesic Distance. Thus, the significance of the assessments based on this model should be interpreted carefully.


## 2.3)
```{r}
printSienaCoev(model.ev)
```
The estimates of the coefficients and the corresponding standard errors are obtained from a model controlling for basic endogenous and exogenous
variables (all variables can be seen in the summary output above). 

H4: Popular students tend to increase or maintain their level of
alcohol consumption

The coefficient for "drinkbeh indegree" is not significantly different from 0 on the 5% significance level. According to our model we do not have evidence that popularity affects the evolution of drinking behavior. Hence, H4 is not supported by our model.

H5: Students tend to adjust their alcohol consumption to that of
their friends.

The coefficient for "drinkbeh average similarity" is significantly different from 0 on the 5% significance level. The coefficient is positive which means that an actor adjust his drinking behavior to the ones of his friends. Hence, H5 is supported.

### Differences for hypotheses 1-3
H1:
The coefficient "indegree - popularity (sqrt)" is not significant any more unlike in the previous model. This means H1 is no longer supported in this model which indicates that the effects are now captured by behavior evolution. 

H2: 
The coefficient "same alcoholConsumption" is now called "same drinkbeh" and is still significant and positive as in the previous model. Thus, H2 is still supported which indicates that the effects are not better captured by the newly included behavior evolution variables.

H3:
The coefficient "logDistance" is still significant and negative as in the previous model. Thus, H3 is still supported.


## 2.4)
According to our model we have evidence that selection and influence processes are occurring. The underlying statistics that support H2 and H3 indicate that there are significant selection processes. Which means we have evidence that actors "select" their friends based on same alcohol consumption (H2) and the distance to them (H3). 

Moreover, the underlying statistic that support H5 indicate that there are significant influence processes too. Meaning there is evidence that students tend to adjust their alcohol consumption to that of their friends. 