---
title: "Assignment 1"
author: "Huang Yingshan, Kahlbacher Fabian, Wattenberg Yannick, Aggeler Samuel"
date: "10/17/2021"
output: html_document
---

# Task 1
## 1)
The probability for observing a tie is equal to the amount of present ties divided by the number of total possible ties.
Let $n = \# \text{of nodes then}$:

$p = \frac{\# \text{of ties}}{\# \text{of possible ties}} = \frac{m}{n \cdot (n-1)} = \frac{174}{1190} \approx 0.146$


## 2)
The probability of observing a specific type of dyad (mutual, asymmetric and null) is the number of dyads of that type divided by the number of total possible dyads.
Let $T = {n \choose 2} = \# \text{of dyads}$:

$p_M = \frac{M}{T} = \frac{52}{595} \approx 0.087$

$p_A = \frac{A}{T} = \frac{70}{595} \approx 0.118$

$p_N = \frac{N}{T} = \frac{473}{595} \approx 0.795$


## 3)
Assuming tie independence with $p = 0.146$ as the probability that a tie exists we get the following values:

$p_M' = P[x_{ij} = 1 \cap x_{ji} = 1] \stackrel{\text{tie ind}.}{=} P[x_{ij} = 1] \cdot P[x_{ji} = 1] = p^2 \approx 0.021$

$p_A' = P[x_{ij} = 1 \cap x_{ji} = 0] + P[x_{ij} = 0 \cap x_{ji} = 1]  \stackrel{\text{tie ind}.}{=}  2 \cdot P[x_{ij} = 1] \cdot P[x_{ji} = 0] = 2 \cdot p \cdot (1-p) \approx 0.25$

$p_N' = P[x_{ij} = 0 \cap x_{ji} = 0] \stackrel{\text{tie ind}.}{=} P[x_{ij} = 0] \cdot P[x_{ji} = 0] = (1 - p)^2 \approx 0.729$


## 4)
The probability of observing mutual dyads in the network is four times higher than what we would expect with tie independence. Further the probability of observing asymmetric dyads
is more than two times higher when assuming tie independence, for null dyads the probability of observation is 6.6\% lower when assuming tie independence.
Given the differences in probability for mutual and asymmetric dyads we would argue that assuming tie independence would not be reasonable.


```{r setup, include=FALSE, echo=TRUE}

set.seed(1908)
library(sna)
library(network)
obsMat <- as.matrix(read.csv("matrix.csv", header=F)) #header needs to be false, look at observed values again, they changed now
cguRec1 <- cug.test(obsMat, grecip, cmode = "edges", reps = 3000)
cguRec2 <- cug.test(obsMat, grecip, cmode = "dyad.census", reps = 3000)
cugTrans <- cug.test(obsMat, gtrans, cmode = "dyad.census",  reps = 3000)
cguRec1
cguRec2
cugTrans
# look for what we're testing for with grecip (testing for both cases at once)
#need to rephrase the hypothesis 

```
## 5)
### 5.1)
The hypothesis of the first test is testing wether there is more reciprocity then density would suggest. The test is conditioned on the edges that means on the network density or number of edges in the network or $U|p$ where p denotes the network density. 

The hypothesis of the second test is the same as for the first test but here we condition on the dyad census. Conditioning on the dyad census here means we condition on the count mutual, asymmetric, and null dyads or $U|MAN$

The hypothesis of the last test is that the amount of transitive relations (ties) that is observed is higher than the amount expected conditioned on the dyad census.

### 5.2)
We think that the second test doesn't make sense because we are testing for the reciprocity of ties conditioned on dyad census. But the dyad census contains the proportion of mutual ties which is equal to the reciprocity of the network by definition. This is also shown in the test results showing we get exactly the value that we observed.

### 5.3)


# Task 2

## 1)

```{r}
set.seed(1908)
library(sna)
library(network)
```

```{r echo=TRUE}
trade2006 <- as.matrix(read.csv("trade1.csv", header=FALSE))
trade2007 <- as.matrix(read.csv("trade2.csv", header=FALSE))
attr <- read.csv("attr.csv")

permutations <- 5000

res_netlm <- netlm(y=trade2007, x=trade2006, reps=permutations, nullhyp="qapy", intercept=T)
#res_netlm <- netlm(y=trade2007, x=trade2006, reps=permutations, nullhyp="qapy", intercept=F)
res_netlm$names <- c("intercept", "trade_06")

summary(res_netlm)

res_nice <- cbind(res_netlm$coefficients, res_netlm$pgreqabs)
colnames(res_nice) <- c("Est.", "p-value")
rownames(res_nice) <- res_netlm$names
res_nice
```
- Should we use an intercept or not??


## 2)
```{r}
# Hp 1
sector <- attr[,2]
same_sector <- outer(sector,sector,"==")*1

# Hp 2
sector_sender_private <- matrix(sector,18,18,byrow=FALSE)
```

## 3)
```{r}
zm <- list(same_sector, sector_sender_private)
res_netlm <- netlm(trade2007, zm, reps=permutations, nullhyp="qapspp")
res_netlm$names <- c("intercept", "same_sector", "sector_sender_private")
summary(res_netlm)
```
The parameter "same_sector" is significantly different from 0 at the significance level $\alpha=0.05$. The parameter indicates that a company which sells goods to another company in the same sector sells 1.1 million CHF less than to another company which is not in the same sector, holding all other variables constant. Thus, the data does not support Hp. 1.

The paramter "sector_sender_private" is not significantly different from 0 at the significance level $\alpha=0.05$. It indicates that companies in the private sector sell 0.14 million CHF more goods than companies which are in the public sector but this relationship is not statistically significant. Therefore, the data does not support Hp 2. 

## 4)
Hp. 3 (size of companies): If the number of employees of a company increases then the amounts of goods sold increases too. 

Hp. 4 (region): The amounts of goods sold by a company is higher if the company which sells is in region A or B than the amount of goods sold by a company which is in region C.

```{r}
# Hp 3.
size_sender <- matrix(attr$size,18,18,byrow=FALSE)

# Hp.4
region_AB <- (attr$region == 1 | attr$region == 2)*1
region_AB_sender <- matrix(region_AB,18,18,byrow=FALSE)
```

## 5)
```{r}
zm <- list(same_sector, sector_sender_private, size_sender, region_AB_sender)
res_netlm <- netlm(trade2007, zm, reps=permutations, nullhyp="qapspp")
res_netlm$names <- c(
  "intercept",
  "same_sector",
  "sector_sender_private",
  "size_sender",
  "region_AB_sender"
)
summary(res_netlm)
```

The parameter "size_sender" is significantly different from 0 at the significance level $\alpha=0.05$. The parameter indicates that for each additional employee the company sells 0.039 million CHF more goods. The data supports Hp 3.

The parameter "region_AB_sender" indicates that companies which operate in the region A or B sell 0.014 million CHF less goods than companies which operate in region C. Since the parameter is not statistically significant from 0 at the significance level $\alpha=0.05$ the data does not support Hp 4.    
